Keep Your Laziness In Check

### 2017-11-20 slides.tex
added images for illustrations
Remove todo about erasing numbers
Add new items to potential questions

\usepackage{pgfpages}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{minted}

@@

\note[itemize]{
\item QuickCheck gives functional correctness
\item But laziness can't be observed by QuickCheck
\item At the 3rd bullet point, say that laziness is not generally considered
      part of functional correctness much like how asymptotic runtime
      is not part of functional correctness because they are not directly
      observable
\item What if we make it observable? Then it's just part of functional correctness!
}

@@

\begin{frame}[fragile]
\frametitle{Observing strictness}
TODO: Just picture?
\begin{minted}{ocaml}
instrumentListWithRef : int ref
                      -> 'a lazyList
                      -> 'a lazyList
\end{minted}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Strictness doesn't exist in a vacuum}
TODO: Pictures with color to show how much is demanded
\end{frame}

\note[itemize]{
  \item These pictures are illustrations of a demand context on a lazy list
  \item These functions have the type from list to unit
  \item It might look like there's only one such function (namely the lazy
        implementation)
  \item There can actually be many different demand behaviors
  \item Forcing the unit value triggers the strictness behavior on the input \texttt{a}
  \item This gives us an easy way to observe the strictness behavior of a context
}

@@

\begin{frame}[fragile]
TODO: erase the numbers in tree
\begin{columns}
\begin{column}{0.7\textwidth}
\frametitle{Beyond lists and numbers}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{binary-tree}
\end{figure}
\end{column}
\begin{column}{0.2\textwidth}
{\fontsize{100}{50}\textrm{\textbf{?}}}
\end{column}
\end{columns}
\end{frame}

@@

\note[itemize]{
  \item Now, we can exactly characterize the demand on a list of Ints by
        composing the type ListDemand and IntDemand
}

@@

\begin{frame}[fragile]
\frametitle{Computing demand, generically}
\begin{itemize}
\item<1->[]
\begin{minted}{ocaml}
demandList : 'b context -> (int lazyList -> 'b)
           -> int lazyList
           -> ('b * intDemand listDemand thunk)

@@

\note[itemize]{
\item There is a determinate relation between a lazy data structure and its
      demand representation
\item We return a \texttt{Thunk} of demand because the input might not be
      demanded at all
\item As the input value is traversed by the function, which is driven by the
      context, the evaluation of each thunk builds a pointer based data
      structure which is isomorphic to the demand data type, which is then
      frozen into a demand representation
\item We expect that the 'a parameter is some kind of lazy structure, and 'b is
      also some kind of lazy structure
\item We use generic programming to implement the DEMAND type for all algebraic
      data types
}

@@

\begin{frame}[fragile]
\frametitle{Connecting specification with observation}
\Huge PLACEHOLDER FOR PICTURE
\end{frame}

\note[itemize]{
  \item To QuickCheck \texttt{takeSpec}, we use mechanisms from QuickCheck to
        fuzz an Int and a [a] as inputs
  \item We can generate random contexts
        that exerts non-trivial demand on the output values
  \item We use the generic \texttt{demand} function to kick off the entire
        computation
  \item We have now observed the demand on the input and also the demand on the
        output of take, so we can just straightforwardly compare that to what
        the specification expects.
  \item If we find a counterexample, we shrink the inputs first, and then shrink
        the reified demand, and re-exert that new demand on the output.
  \item Higher-order also works! but we don't want to go into the details
}

\iffalse
\begin{frame}[fragile]
\frametitle{Higher-order specifications}
\Huge PLACEHOLDER FOR PICTURE
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Have questions?}
Come ask us about \dots
\begin{itemize}
\item Higher-order specifications
\item Generic programming
\item Random generation of demand contexts
\item Details of instrumentation
\item What language are you actually implementing this in? \\ (Hint: it's not ML)
\item[] \dots
\end{itemize}
\end{frame}

>>>>

\usepackage{pgfpages}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{float}
\usepackage{minted}

@@

\note[itemize]{
\item PBRT is a technique for randomized testing which allows users to write
  down an executable specification for a function and test whether it holds for
  randomly generated inputs.
\item QuickCheck gives functional correctness
\item But laziness can't be observed by QuickCheck
\item Say that laziness is not generally considered
      part of functional correctness much like how asymptotic runtime
      is not part of functional correctness because they are not directly
      observable
\item What if we make it observable? Then it's just part of functional correctness!
}

@@

\begin{frame}[fragile]
\frametitle{Observing strictness}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{instrumentListWithRef}
\caption{Instrumented List}
\end{figure}
\begin{minted}{ocaml}
instrumentListWithRef : int ref
                      -> 'a lazyList
                      -> 'a lazyList
\end{minted}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Strictness doesn't exist in a vacuum}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{noDemand}
\caption{No Demand}
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Strictness doesn't exist in a vacuum}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{whnf}
\caption{Weak Head Normal Form}
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Strictness doesn't exist in a vacuum}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{spinestrict}
\caption{Spine Strict}
\end{figure}
\end{frame}

\note[itemize]{
  \item These pictures are illustrations of a demand context on a lazy list
  \item This gives us an easy way to observe the strictness behavior of a
    context
  \item While these pictures might look like a data structure, they are meant to
    describe the behavior of a context of evaluation on a lazy data structure of
    this shape }

@@

\begin{frame}[fragile]
\begin{columns}
\begin{column}{0.7\textwidth}
\frametitle{Beyond lists and numbers}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{binary-tree}
\end{figure}
\end{column}
\begin{column}{0.2\textwidth}
{\fontsize{100}{50}\textrm{\textbf{?}}}
\end{column}
\end{columns}
\end{frame}

@@

\note[itemize]{
\item Now, we can exactly characterize a demand on a list of Ints by composing
  the type ListDemand and IntDemand
\item The type ListDemand is a reification of the demand placed upon a lazy list
  during the course of evaluation in some context }

@@

\begin{frame}[fragile]
\frametitle{Observing demand, generically}
\begin{itemize}
\item<1->[]
\begin{minted}{ocaml}
demandList : 'b context -> (int lazyList -> 'b)
           -> int lazyList
           -> ('b * intDemand listDemand thunk)

@@

\note[itemize]{
\item There is a determinate relation between a lazy data structure and its
      demand representation
\item We return a \texttt{Thunk} of demand because the input might not be
      demanded at all
\item We expect that the 'a parameter is some kind of lazy structure, and 'b is
      also some kind of lazy structure
\item We use generic programming to implement the DEMAND type for all algebraic
      data types
}

@@

\begin{frame}[fragile]
\frametitle{Connecting specification with observation}
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{specflow}
\end{figure}
\end{frame}

\note[itemize]{
\item We randomly fuzz the inputs to the function, and
  \item We can generate random contexts
        that exert non-trivial demand on the output values
  \item We use the generic \texttt{demand} function to kick off the entire
        computation
  \item We have now observed the demand on the input and also the demand on the
        output, so we can just straightforwardly compare that to what
        the specification expects.
  \item If we find a counterexample, we shrink the inputs first, and then shrink
        the reified demand, and re-exert that new demand on the output.
  \item Higher-order also works! but we don't want to go into the details
}

\iffalse
\begin{frame}[fragile]
\frametitle{Higher-order specifications}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Have questions?}
Come ask us about \dots
\begin{itemize}
\item Higher-order specifications
\item Generic programming for any arity and datatype
\item Random generation of partial demand contexts
\item Efficiency of our implementation
\item Details of instrumentation
\item What language are you actually implementing this in? \\ (Hint: it's not ML)
\item[] \dots
\end{itemize}
\end{frame}
